```{r echo=FALSE}
options(width=100)
knitr::opts_knit$set(verbose = TRUE)
```

Quality control of HiSeq runs before alignment
==============================================

Summary
-------

 - Yield is comparable between runs.

 - There is an obvious defect, however, which is the mixing of rows `08` and `09`
   in run `1772-064-103`.  This is indicated by a `FALSE` value in the column
   `HiSeq_QC`.

 - Libraries with less than 100,000 pairs are also very likely to be
   problematic and are flagged `FALSE` in `HiSeq_QC`.


Datasets
--------

The `Demultiplex_Stats.htm` files from the _Basecall Stats_ directory of the
HiSeq runs were copied and renamed with their run ID, opened in a web browser,
and the table inside was pasted in LibreOffice, edited and saved in _xslx_
format before loading into `R`.

 - `1772-067-038` and `1772-067-039` C1 runs: HiSeq run ID [140305_700213F_0043_BH0TMJADXX](./140305_700213F_0043_BH0TMJADXX.html).

 - `1772-064-103` C1 run: HiSeq run ID [140124_SN7001394_0068_AH0TK5ADXX](./140124_SN7001394_0068_AH0TK5ADXX.html).

 - `1772-062-248` and `1772-062-249` C1 runs: HiSeq run ID [131205_SN7001394_0065_BH0BG5ADXX](./131205_SN7001394_0065_BH0BG5ADXX.html).

The edition in LibreOffice was as follows:

 - Added a `Run` column, with the C1 run ID.

 - Renamed the `Sample ID` value for the _Undetermined_ sequences `Undetermined`.
   Example: `lane1` renamed to `Undetermined`.

 - For the same lines, renamed the `Description` value to our facility's internal ID.
   Example: `Clusters with unmatched barcodes for lane 1 ` renamed to `LS2079`.

 - For 1772-062-248 and 1772-062-249, entered the `Sample ID` by hand, using
   same indexes as for the libraries sequenced later.

The files were then collated in a single Excel file, `HiSeq.xslx`.


Quality control
---------------

```{r message=FALSE}
library(gdata)
library(ggplot2)
```

### Load data

```{r}
qc <- read.xls( "HiSeq.xlsx",
  col.names=c(
    'Run',
    'Lane',
    'Sample.ID',
    'Sample.Ref',
    'Index',
    'Description',
    'Control',
    'Project',
    'Yield',
    'Percent.PF',
    'Reads',
    'raw.clusters.per.lane',
    'Perfect.Index.Reads',
    'One.Mismatch.Index.Reads',
    'Q30.bases',
    'Mean.Quality.Score'))
```

The `Sample.ID` column contains identifiers made of the library ID attributed
by our facility, followed by `-`, follow by a well ID in 96-well plate format.

`row` and `column` refer to the coordinates in the 96-well plates (where the multiplexing reaction
was done).

```{r}
qc$Lane <- factor(qc$Lane)
qc$Well <- factor(sub('RNhi.....-', '', qc$Sample.ID))
qc[qc$Well == 'Undetermined', 'Well'] <- NA
qc$row    <- factor(substr(qc$Well, 1, 1))
qc$column <- factor(substr(qc$Well, 2, 3))
qc$Q30.bases <- as.numeric(as.character(qc$Q30.bases))
qc$Mean.Quality.Score <- as.numeric(as.character(qc$Mean.Quality.Score))
qc$Library <- sub('-...','', qc$Sample.ID)
qc$Library <- factor(qc$Library)

summary(qc)
```


### Quantity of demultiplexed and unmatched paired-end reads.

The report files also provide information on non-demultiplexed reads, in lines
where `Sample.ID` equals `Undetermined`.  This creates `NA` values in the
`Well` column.

```{r}
Unmatched <- factor(is.na(qc$Well), labels=c('Demultiplexed', 'Unmatched'))
tapply(qc$Reads, list(qc$Run, Unmatched), sum) / 2
```

The numbers are divided by two because the number of reads reported is
paired-end (two reads per pair).


### Looking for run bias

```{r HiSeq_run_comparison, dev='svg'}
qplot(data = qc[!is.na(qc$Well),], Run, Reads, geom = "boxplot")
```

### Looking for position bias in the plates

Do not worry too much about the shape of the boxplots, this is because there
are few points per run (8 for columns and 12 for rows).

```{r HiSeq_plate_comparison, dev='svg'}
qplot(data = qc[!is.na(qc$Well),], row, Reads, geom = "boxplot") + facet_wrap(~Run)
qplot(data = qc[!is.na(qc$Well),], column, Reads, geom = "boxplot") + facet_wrap(~Run)
```

In C1 run `1772-064-103`, the plate column `09` has clearly been pipetted in
row `08`.  Both of them are flagged `FALSE` in a table column called `HiSeq_QC`
to mark that fact.

```{r}
qc$HiSeq_QC <- TRUE
qc[qc$Run == '1772-064-103' & grepl('0[89]', qc$Well), 'HiSeq_QC'] <- FALSE
```

### Libraries with less than 100,000 pairs.

The libraries with less than 200,000 reads (one pair is two reads) are likely
to be problematic or too shallow in comparison with the others, and therefore
flagged `FALSE` for the quality control.

```{r}
qc[qc$Reads < 200000, 'HiSeq_QC'] <- FALSE
```

Export table in CSV format.
---------------------------

Drop columns that contain the same value for each library.

```{r}
qc <- qc[,-grep('One.Mismatch.Index.Reads', colnames(qc))]
qc <- qc[,-grep('Control', colnames(qc))]
```

Rename `Description` to `LSID` to avoid name conflicts with the fluorescence data.

```{r}
colnames(qc) <- sub('Description', 'LSID', colnames(qc))
```

Write the table in CSV format.

```{r}
qc$cell_id <- paste(qc$Run, qc$Well, sep = "_")
summary(qc)
write.csv(file='HiSeq.csv', qc, row.names=FALSE)
```
