```{r echo=FALSE}
knitr::opts_knit$set(verbose = TRUE)
options(width=100)
```

Fluorescence data correction
=============================

The input is a data frame with 6 columns: "cell_id", "mean_ch2", "bg_mean_ch2", "mean_ch3", "bg_mean_ch3", "Run". The data file is `Results_fluorescence.csv`.

Install all the packages listed below with `install.packages()`, in case they are not already installed. Then load the packages.

```{r load required libraries}
library(moments)
library(lattice)
library(flexmix)
library(limma)
library(MASS)
library(rms)
library(contrast)
```

```{r function collection}
# function for creating a list of input data objects
createData <- function(data) {
  dd <- matrix(as.numeric(data[, 2:(ncol(data)-1)]), nrow=nrow(data))
  samples <- data[,1]
  # assigns a number to all 5 runs
  batches <- as.numeric(data[,ncol(data)])
  # creates a list of 4 elements (row index for each cell, matrix of fluorescence values, sample name factor, batch number)
  return(list(index=1:length(samples), RGexprs=dd, samples=samples, batch=batches))
}

BGcorrectFucci <- function(data, method, old.offset, bg) {
  if(bg == TRUE) {
    RG <- new("RGList", list(R=data[,1], G=data[,3], Rb=data[,2], Gb=data[,4]))
    offset <- min(apply(data[, c(2,4)], 2, min))+1
  if(method == "subtract") {
    core<-backgroundCorrect(RG, method=method, offset=offset)
  } 
  else {
    core <- backgroundCorrect(RG, method=method)
  }}
  if(bg == FALSE) {
    RG <- new("RGList", list(R=data[,1], G=data[,2]))
    core <- backgroundCorrect(RG, method=method, offset=old.offset)
  }
  core<-matrix(cbind(core$R, core$G), ncol=2)
  return(list(core, offset))
}

boxcoxMatrix <- function(data) {
  est <- boxcox(data~1, lambda=seq(-4, 4, 0.01), plotit=FALSE)
  lmb <- est$x[which(est$y == max(est$y))]
  rang <- range(est$x[est$y > max(est$y)-qchisq(0.95, 1)/2])
  if(rang[1]*rang[2]>=0) {
  data <- log(data)
  lmb <- 0
  } else {
  data <- ((data^lmb)-1)/lmb
  }
  return(list(data, lmb))
}


doTransform <- function(data, transform) {
  dat <- c()
  if(transform == "bc") {
  est <- apply(data, 2, boxcoxMatrix)
  dat <- matrix(cbind(matrix(est[[1]][[1]], ncol=1), matrix(est[[2]][[1]], ncol=1)), ncol=2)
  lpar <- c(est[[1]][[2]], est[[2]][[2]])
  }
  if(transform == "log") {
  dat <- matrix(log(data), ncol=2)
  lpar <- c()
  }
  if(transform == "log10") {
  dat <- matrix(log(data, 10), ncol=2)
  lpar <- c()
  }
  if(transform == "asinh") {
  dat <- matrix(asinh(data), ncol=2)
  lpar <- c()
  }
  if(transform == "none") {
  dat <- data
  lpar <- c()
  }
  return(list(dat, lpar))
}


invTransform <- function(data, lambda, transform) {
  res <- c()
  if(transform == "bc") {
  if(lambda == 0) { 
  res <- exp(data) 
  } else {
  res <- (lambda*data + 1)^(1/lambda)
  }}
  if(transform == "log10") {
  res <- 10^data
  }
  if(transform == "log") {
  res <- exp(data)
  }
  if(transform == "asinh") {
  res <- sinh(data)
  }
  return(res)
} 


boxFucci <- function(data, transform, reference, legends) {
  edata <- data$exprs
  f <- factor(data$batch)
  f <- as.numeric(levels(relevel(f, ref=reference)))
  d <- as.list(rep(0, (2*max(f))))
  est <- doTransform(data=edata, transform=transform)
  dat <- est[[1]]
  lpar <- est[[2]]
  d[[f[1]]] <- dat[which(data$batch == f[1]), 1]
  for(i in 2:(2*max(f))) {
  if(i<=max(f)) {
  d[[f[i]]] <- dat[which(data$batch == f[i]), 1]
  } else {
  d[[(f[(i-max(f))]+max(f))]] <- dat[which(data$batch == f[(i-max(f))]), 2]
  }}
  
  legs <- c(paste("Ch2: ", max(f), " Runs", sep=""), paste("Ch3: ", max(f), " Runs", sep=""))
  densX <- as.list(rep(0, length(d)))
  densY <- as.list(rep(0, length(d)))
  for(i in 1:length(d)) {
  densX[[i]] <- density(d[[i]])$x
  densY[[i]] <- density(d[[i]])$y
  }
  adX <- unlist(densX)
  adY <- unlist(densY)
  
  plot(densX[[1]], densY[[1]], main=legs[1], col=1, type="l", sub="", xlab="transformed intensities", ylab="Density", xlim=c(min(adX), max(adX)), ylim=c(min(adY), max(adY)))
  if(max(f)>1) {
  for(i in 2:max(f)) {
  lines(densX[[i]], densY[[i]], col=i)
  }}
  
  plot(densX[[(1+max(f))]], densY[[(1+max(f))]], main=legs[2], col=1, type="l", sub="", xlab="transformed intensities", ylab="Density", xlim=c(min(adX), max(adX)), ylim=c(min(adY), max(adY)))
  if(max(f)>1) {
  for(i in 2:max(f)) {
  lines(densX[[(i+max(f))]], densY[[(i+max(f))]], col=i)
  }}
  
  return(list(dat, data$batch, lpar))
}


refineMixes <- function(data, batch, model) {
  cc1 <- clusters(model)
  cc <- rep(0, length(cc1))
  for(i in 1:max(batch)) {
  w <- which(batch == i)
  cl <- cc1[w]
  d <- data[w]
  sl <- sort.list(aggregate(d, list(cl), mean)[,2])
  cl1 <- rep(0, length(cl))
  for(j in 1:length(sl)) {
  cl1[which(cl == sl[j])] <- j
  }
  cc[w] <- cl1
  }
  return(cc)
}


lmFucci <- function(data, batch, maxMix, reference, prior.pi) {
  if(length(unique(batch)) == 1) {
  Batch <- factor(batch)
  mod <- stepFlexmix(data ~ 1, k = 1:maxMix, nrep=20, control = list(minprior = prior.pi))
  }
  if(length(unique(batch))>1) {
  Batch <- factor(batch)
  Batch <- relevel(Batch, ref=reference)
  mod <- stepFlexmix(data ~ Batch, k = 1:maxMix, nrep=20, control = list(minprior = prior.pi))
  }
  gg <- getModel(mod)
  Comp <- refineMixes(data, batch, gg)
  w1 <- c()
  
  if(length(unique(batch))>1 & length(unique(Comp))>1) {
  mod <- lm(data~factor(Comp)*factor(Batch))
  desMat <- model.matrix(~factor(Comp)*factor(Batch))
  rr <- rownames(summary(mod)[[4]])
  w1 <- which(rr == paste("factor(Comp)", max(Comp), sep=""))
  }
  if(length(unique(batch))>1 & length(unique(Comp)) == 1) {
  mod <- lm(data~factor(Batch))
  desMat <- model.matrix(~factor(Batch))
  rr <- rownames(summary(mod)[[4]])
  w1 <- which(rr == "(Intercept)")
  }
  if(length(unique(batch)) == 1 & length(unique(Comp))>1) {
  mod <- lm(data~factor(Comp))
  desMat <- model.matrix(~factor(Comp))
  }
  if(length(unique(batch)) == 1 & length(unique(Comp)) == 1) {
  mod <- lm(data~1)
  desMat <- c()
  }
  rr <- rownames(summary(mod)[[4]])
  cc <- c(colnames(summary(mod)[[4]]), "FDR")
  compEsts <- matrix(as.numeric(summary(mod)[[4]][,c(1,4)]), ncol=2)
  compEsts <- matrix(cbind(compEsts, matrix(p.adjust(compEsts[, ncol(compEsts)], "BH"), ncol=1)), nrow=nrow(compEsts))
  resids <- mod$residuals
  stdresids <- rstandard(mod)
  fitted <- mod$fitted.values
  compEsts <- matrix(cbind(rr, compEsts), ncol=(ncol(compEsts)+1))
  compEsts <- matrix(rbind(c("", cc[c(1, 4, 5)]), compEsts), ncol=ncol(compEsts))
  ww <- as.list(rep(0, max(Comp)))
  for(i in 1:max(Comp)) {
  ww[[i]] <- which(Comp == i)
  }
  wall <- length(rr)
  subtInd <- c()
  if(length(w1)>0) {
  subtInd <- (w1+2):(wall+1)
  }
  return(list(compEsts, Comp, ww, resids, stdresids, fitted, subtInd, desMat))
}


batchFucci  <-  function(data, maxMix, reference, prior.pi) {
  edata <- data[[1]]
  om <- apply(edata, 2, mean)
  ests <- apply(edata, 2, lmFucci, batch=data[[2]], maxMix=maxMix, reference=reference, prior.pi=prior.pi)
  coefs <- list(as.numeric(ests[[1]][[1]][ests[[1]][[7]], 2]), as.numeric(ests[[2]][[1]][ests[[2]][[7]], 2]))
  new <- resids <- stdresids <- fitted <- edata
  for(i in 1:2) {
  resids[, i] <- as.numeric(ests[[i]][[4]])
  stdresids[, i] <- as.numeric(ests[[i]][[5]])
  fitted[, i] <- as.numeric(ests[[i]][[6]])
  if(length(ests[[i]][[7]])>0) {
  for(b in 1:length(ests[[i]][[7]])) {
  new[, i] <- new[, i]-coefs[[i]][b]*ests[[i]][[8]][, (ests[[i]][[7]][b]-1)]
  }
  }
  new[, i] <- new[, i]-mean(new[, i])+om[i]
  }
  return(list(exprs=new, batch=data[[2]], mixesCh2=ests[[1]][[2]], mixesCh3=ests[[2]][[2]], residuals=resids, standardized.residuals=stdresids, fitted.values=fitted, estCh2=ests[[1]][[1]], estCh3=ests[[2]][[1]], designCh2=ests[[1]][[8]], designCh3=ests[[2]][[8]]))
}


# the main fuction to create necessary parameters for normaization
# corresponds to step 2
adjustFucci <- function(data, transform, qnormalize=FALSE, maxMix, reference, by.batch=FALSE, prior.pi, savePlot) {
  if(length(savePlot)>0) {
  postscript(paste(savePlot, "/densities.ps", sep=""))
  }
  par(mfrow=c(2, 2))
  if(by.batch == FALSE) {
  dd <- BGcorrectFucci(data$RGexprs, method="subtract", old.offset=c(), bg=TRUE)
  dd <- list(exprs=dd[[1]], batch=data$batch, offset=dd[[2]])
  if(reference>max(data$batch)) {
  reference <- max(data$batch)
  print(paste("the baseline dataset for batch effect correction has changed to ", reference, sep=""))
  }
  res1 <- boxFucci(dd, transform=transform, reference=reference, legends="uncorrected/unadjusted")
  res2 <- batchFucci(res1, maxMix=maxMix, reference=reference, prior.pi=prior.pi)
  res2$exprs <- BGcorrectFucci(matrix(cbind(invTransform(res2$exprs[, 1], lambda=res1[[3]][1], transform=transform), invTransform(res2$exprs[, 2], lambda=res1[[3]][2], transform=transform)), ncol=2), method="normexp", old.offset=dd$offset, bg=FALSE)[[1]]
  res3 <- boxFucci(res2, transform=transform, reference=reference, legends="corrected/unadjusted")
  } else {
  d1 <- matrix(0, 1, 2)
  for(b in 1:max(data$batch)) {
  d1 <- matrix(rbind(d1, BGcorrectFucci(data$RGexprs[which(data$batch == b), ], method="normexp", old.offset=c(), bg=TRUE)[[1]]), ncol=2)
  }
  dd <- d1[-1, ]
  dd <- list(exprs=dd, batch=data$batch)
  if(reference>max(data$batch)) {
  reference <- max(data$batch)
  print(paste("the baseline dataset for batch effect correction has changed to ", reference, sep=""))
  }
  res1 <- boxFucci(dd, transform=transform, reference=reference, legends="uncorrected/unadjusted")
  res2 <- batchFuccii(res1, maxMix=maxMix, reference=reference, prior.pi=prior.pi)
  res2$exprs <- matrix(cbind(invTransform(res2$exprs[, 1], lambda=res1[[3]][1], transform=transform), invTransform(res2$exprs[, 2], lambda=res1[[3]][2], transform=transform)), ncol=2)
  res3 <- boxFucci(res2, transform=transform, reference=reference, legends="corrected/unadjusted")
  }
  if(qnormalize == TRUE) {
  res3[[1]] <- normalize.quantiles(res3[[1]])
  res3 <- boxFucci(res3[[1]], transform="none", reference=reference, legends="corrected/adjusted")
  }
  
  if(length(savePlot)>0) {
  dev.off()
  }
  
  if(length(savePlot) == 0) {
  x11()
  }
  r <- res2$residuals
  sr <- res2$standardized.residuals
  f <- res2$fitted.values
  
  if(length(savePlot)>0) {
  postscript(paste(savePlot, "/diagnostics.ps", sep=""))
  }
  par(mfrow=c(3, 3))
  options(warn=-1)
  hist(c(r), breaks=50, xlab="Residuals", main="Model Residuals", sub=paste("KS-test for normality: ", round(ks.test(c(r), "pnorm", 0, sqrt(var(c(r))))$p.value, 3), sep=""))
  hist(r[, 1], breaks=50, xlab="Residuals", main="Residuals of Ch2 Fucci", sub=paste("KS-test for normality: ", round(ks.test(r[, 1], "pnorm", 0, sqrt(var(r[, 1])))$p.value, 3), sep=""))
  hist(r[, 2], breaks=50, xlab="Residuals", main="Residuals of Ch3 Fucci", sub=paste("KS-test for normality: ", round(ks.test(r[, 2], "pnorm", 0, sqrt(var(r[, 2])))$p.value, 3), sep=""))
  plot(c(f), c(sr), xlab="Model fitted values", ylab="Model Standardized Residuals", main="Fitted vs standardized residuals (Model)")
  plot(f[, 1], sr[, 1], xlab="Ch2 fitted values", ylab="Ch2 Standardized Residuals", main="Fitted vs standardized residuals (Ch2)")
  plot(f[, 2], sr[, 2], xlab="Ch3 fitted values", ylab="Ch3 Standardized Residuals", main="Fitted vs standardized residuals (Ch3)")
  cr <- c(r)
  cf <- c(f)
  acf(cr[sort.list(cf)], main="Autocorrelation of Model residuals")
  acf(r[sort.list(f[, 1]), 1], main="Autocorrelation of Ch2 residuals")
  acf(r[sort.list(f[, 2]), 2], main="Autocorrelation of Ch3 residuals")
  
  if(length(savePlot)>0) {
  dev.off()
  }
  if(length(savePlot) == 0) {
  x11()
  }
  if(length(savePlot) > 0) {
  postscript(paste(savePlot, "/corrected2D.ps", sep=""))
  }
  plot(res3[[1]], cex = 0.8, xlab="Ch2 (green channel), corrected intensity", ylab="Ch3 (red channel), corrected intensity", main="")
  if(length(savePlot) > 0) {
  dev.off()
  }
  
  ago <- c(round(agostino.test(r[, 1])$p.value, 3), round(agostino.test(r[, 2])$p.value, 3), round(agostino.test(c(r))$p.value, 3))
  bon <- c(round(bonett.test(r[, 1])$p.value, 3), round(bonett.test(r[, 2])$p.value, 3), round(bonett.test(c(r))$p.value, 3))
  jar <- c(round(jarque.test(r[, 1])$p.value, 3), round(jarque.test(r[, 2])$p.value, 3), round(jarque.test(c(r))$p.value, 3))
  ks <- c(round(ks.test(r[, 1], "pnorm", 0, sqrt(var(r[, 1])))$p.value, 3), round(ks.test(r[, 2], "pnorm", 0, sqrt(var(r[, 2])))$p.value, 3), round(ks.test(c(r), "pnorm", 0, sqrt(var(c(r))))$p.value, 3))
  sk <- c(round(skewness(r[, 1]), 3), round(skewness(r[, 2]), 3), round(skewness(c(r)), 3))
  ku <- c(round(kurtosis(r[, 1]), 3), round(kurtosis(r[, 2]), 3), round(kurtosis(c(r)), 3))
  legR <- c("", "Skewness (ideal=0)", "Kurtosis (ideal=3)", "Agostino test for Skewness", "Bonett test for Kurtosis", "Jarque test for Normality", "KS-test for Normality")
  legC <- c("Ch2", "Ch3", "Ch2 & Ch3") 
  report <- matrix(cbind(legC, sk, ku, ago, bon, jar, ks), nrow=3)
  report <- matrix(rbind(legR, report), ncol=ncol(report))
  resALL <- c(data, list(exprs=dd$exprs, corrected.exprs=res2$exprs, corrected.transformed.exprs=res3[[1]], mixesCh2=res2$mixesCh2, mixesCh3=res2$mixesCh3, BatchCh2.est=res2$estCh2, BatchCh3.est=res2$estCh3, fitted.values=f, transform=transform, model.residuals=r, model.standardized.residuals=sr, residual.statistics=report, lpar=res1[[3]], designCh2=res2$designCh2, designCh3=res2$designCh3, reference=reference))
  return(resALL)
}


# calculates contrast estimates (differences and P-values) among the runs
# corresponds to step 3
contrastFucci <- function(data, channel) {
  data$exprs <- doTransform(data$exprs, data$transform)[[1]]
  if(channel == "Ch2") {
  d1 <- data.frame(Int=data$exprs[, 1], mixes=factor(data$mixesCh2), batch=relevel(factor(data$batch), ref=data$reference))
  a <- as.character(sort(unique(data$mixesCh2)))
  }
  if(channel == "Ch3") {
  d1 <- data.frame(Int=data$exprs[, 2], mixes=factor(data$mixesCh3), batch=relevel(factor(data$batch), ref=data$reference))
  a <- as.character(sort(unique(data$mixesCh3)))
  }
  mod <- lm(Int~mixes*batch, data=d1)
  b <- c()
  for(i in 1:max(data$batch)) {
  b <- c(b, as.character(i))
  }
  b <- t(combn(b, 2))
  res <- matrix(c("Channel", "Component", "Run1", "Run2", "Contrast", "Pvalue"), 1, 6)
  for(i in 1:length(a)) {
  for(j in 1:nrow(b)) {
  rr <- contrast(mod,  a=list(mixes=a[i], batch=b[j, 1]),  b=list(mixes=a[i], batch=b[j, 2]),  type = "average")
  res <- matrix(rbind(res, c(channel, a[i], b[j, ], as.numeric(unlist(rr)[c(1, 7)]))), ncol=ncol(res))
  }}
  
  p <- p.adjust(as.numeric(res[2:nrow(res), ncol(res)]), "BH")
  res <- matrix(cbind(res, c("FDR", p)), nrow=nrow(res))
  ll <- list(res)
  names(ll) <- paste(channel, ".contrasts", sep="")
  return(c(data, ll))
}

```

The following part uses the Results_fluorescence.csv file generated by the Fluorescence-measured-in-ImageJ.Rmd script.
Step 1 brings the data to a required input format for the background and run effects correction, which is done in step 2.

```{r final intensities, dev=c("svg", "png")}
# subsets and transforms the input data
data <- read.csv("../fluorescence/Results_fluorescence.csv", header=T)
data <- data[!data$fluo_QC, c("cell_id", "mean_ch2", "bg_mean_ch2", "mean_ch3", "bg_mean_ch3", "Run")]
data$Run <- as.numeric(data$Run)
data <- as.matrix(data)

step1 <- createData(data)
step2 <- adjustFucci(data=step1, transform="bc", maxMix=3, reference=5, prior.pi=0.2, savePlot=c())
step3 <- contrastFucci(data=step2, channel="Ch2")
final <- contrastFucci(data=step3, channel="Ch3")
# create table with corrected fluorescence values
correctedFluo <- data.frame(final$samples, final$batch, final$RGexprs[,1:4], exp(final$exprs), final$corrected.exprs, final$corrected.transformed.exprs, final$mixesCh2, final$mixesCh3, final$model.residuals)
names(correctedFluo) <- c("cell_id", "batch", "Ch2_foreground", "Ch2_background", "Ch3_foreground", "Ch3_background",
"Ch2_unadjusted", "Ch3_unadjusted", "ch2_corrected", "ch3_corrected", "ch2_corrected_transformed", "ch3_corrected_transformed", "Ch2_mix_components", "Ch3_mix_components", "Ch2_flexmix_residuals", "Ch3_flexmix_residuals")
write.csv(correctedFluo, "correctedIntensities.csv", row.names=F)
```

The last section uses the data with a simplified model and compares the result with the flexmix model results above.

```{r testing of alternative model}
# get the data of interest
x <- final$RGexprs[, 1:4]
x <- matrix(cbind(apply(x[, 2:1], 1, diff), apply(x[, 4:3], 1, diff), final$batch), ncol=3)

# puts a small positive value to negative signals (to avoid missing data because of log-transform below)
# this resembles what normexp is supposed to do
 
x[x<=0] <- 0.00001
l1 <- lm(log(x[, 1])~factor(x[, 3]))
l2 <- lm(log(x[, 2])~factor(x[, 3]))

l1c <- l1$coefficients
l2c <- l1$coefficients
corr1 <- log(x[, 1])
mod1 <- model.matrix(~factor(x[, 3]))
for(i in 2:5){
corr1 <- corr1-l1c[i]*mod1[, i]
}

corr2 <- log(x[, 2])
mod2 <- model.matrix(~factor(x[, 3]))
for(i in 2:5){
corr2 <- corr2-l2c[i]*mod2[, i]
}

signals <- matrix(cbind(corr1, corr2), ncol=2)
resids <- matrix(cbind(l1$residuals, l2$residuals), ncol=2)

#join the data
correctedFluo <- data.frame(correctedFluo, Ch2_corrected_transformed_anova=signals[, 1], 
Ch3_corrected_transformed_anova=signals[, 2], Ch2_residuals_anova=resids[, 1], 
Ch3_residuals_anova=resids[, 2])
```

```{r density and residual histogram plots}
# the data of interest
x <- correctedFluo

pdf(file="densities.pdf")

yscale <- density(x$Ch2_unadjusted[which(x$batch==1)])
plot(density(x$Ch2_unadjusted[which(x$batch==1)]), ylim=c(0, max(yscale$y)), col=1, xlab="Unadjusted log-signal", main="Figure A: Densities of unadjusted Ch2 signals")
lines(density(x$Ch2_unadjusted[which(x$batch==2)]), col=2)
lines(density(x$Ch2_unadjusted[which(x$batch==3)]), col=3)
lines(density(x$Ch2_unadjusted[which(x$batch==4)]), col=4)
lines(density(x$Ch2_unadjusted[which(x$batch==5)]), col=5)
legend("topright", title="Ch2 signals", c("1772-062-248", "1772-062-249", "1772-064-103", "1772-067-038", "1772-067-039"), lty=1, col=1:5, cex=0.75)

plot(density(x$Ch3_unadjusted[which(x$batch==1)]), ylim=c(0, 1), xlab="Unadjusted log-signal", col=1, main="Figure B: Densities of unadjusted Ch3 signals")
lines(density(x$Ch3_unadjusted[which(x$batch==2)]), col=2)
lines(density(x$Ch3_unadjusted[which(x$batch==3)]), col=3)
lines(density(x$Ch3_unadjusted[which(x$batch==4)]), col=4)
lines(density(x$Ch3_unadjusted[which(x$batch==5)]), col=5)
legend("topleft", title="Ch3 signals", c("1772-062-248", "1772-062-249", "1772-064-103", "1772-067-038", "1772-067-039"), lty=1, col=1:5, cex=0.75)

plot(density(x$ch2_corrected_transformed[which(x$batch==1)]), col=1, ylim=c(0, 1), xlab="Adjusted log-signal", main="Figure C: Densities of adjusted Ch2 signals (Flexmix)")
lines(density(x$ch2_corrected_transformed[which(x$batch==2)]), col=2)
lines(density(x$ch2_corrected_transformed[which(x$batch==3)]), col=3)
lines(density(x$ch2_corrected_transformed[which(x$batch==4)]), col=4)
lines(density(x$ch2_corrected_transformed[which(x$batch==5)]), col=5)
legend("topright", title="Ch2 signals", c("1772-062-248", "1772-062-249", "1772-064-103", "1772-067-038", "1772-067-039"), lty=1, col=1:5, cex=0.75)

plot(density(x$ch3_corrected_transformed[which(x$batch==1)]), col=1, ylim=c(0, 1), xlab="Adjusted log-signal", main="Figure D: Densities of adjusted Ch3 signals (Flexmix)")
lines(density(x$ch3_corrected_transformed[which(x$batch==2)]), col=2)
lines(density(x$ch3_corrected_transformed[which(x$batch==3)]), col=3)
lines(density(x$ch3_corrected_transformed[which(x$batch==4)]), col=4)
lines(density(x$ch3_corrected_transformed[which(x$batch==5)]), col=5)
legend("topright", title="Ch3 signals", c("1772-062-248", "1772-062-249", "1772-064-103", "1772-067-038", "1772-067-039"), lty=1, col=1:5, cex=0.75)

plot(density(x$Ch2_corrected_transformed_anova[which(x$batch==1)]), col=1, ylim=c(0, 0.6), xlab="Adjusted log-signal", main="Figure E: Densities of adjusted Ch2 signals (1-way ANOVA)")
lines(density(x$Ch2_corrected_transformed_anova[which(x$batch==2)]), col=2)
lines(density(x$Ch2_corrected_transformed_anova[which(x$batch==3)]), col=3)
lines(density(x$Ch2_corrected_transformed_anova[which(x$batch==4)]), col=4)
lines(density(x$Ch2_corrected_transformed_anova[which(x$batch==5)]), col=5)
legend("topright", title="Ch2 signals", c("1772-062-248", "1772-062-249", "1772-064-103", "1772-067-038", "1772-067-039"), lty=1, col=1:5, cex=0.75)

plot(density(x$Ch3_corrected_transformed_anova[which(x$batch==1)]), col=1, ylim=c(0, 0.6), xlab="Adjusted log-signal", main="Figure F: Densities of adjusted Ch3 signals (1-way ANOVA)")
lines(density(x$Ch3_corrected_transformed_anova[which(x$batch==2)]), col=2)
lines(density(x$Ch3_corrected_transformed_anova[which(x$batch==3)]), col=3)
lines(density(x$Ch3_corrected_transformed_anova[which(x$batch==4)]), col=4)
lines(density(x$Ch3_corrected_transformed_anova[which(x$batch==5)]), col=5)
legend("topright", title="Ch3 signals", c("1772-062-248", "1772-062-249", "1772-064-103", "1772-067-038", "1772-067-039"), lty=1, col=1:5, cex=0.75)


#residual histograms}
r1 <- c(x$Ch2_residuals_anova)
ks1 <- round(ks.test(r1, "pnorm", 0, sqrt(var(r1)))$p.value, 6)
hist(r1, br=110, main="Figure G: Residuals of 1-way ANOVA adjustment", xlab="Residuals", sub=paste("KS-test: ", ks1, sep=""), xlim=c(-15,5))

r2 <- c(x$Ch2_flexmix_residuals)
ks2 <- round(ks.test(r2, "pnorm", 0, sqrt(var(r2)))$p.value, 6)
hist(r2, br=110, main="Figure H: Residuals of Flexmix adjustment", xlab="Residuals", sub=paste("KS-test: ", ks2, sep=""), xlim=c(-1,1)) 

dev.off()
```